---
title: "Gillespie - Project1"
author: "Scott Gillespie"
date: "April 11, 2016"
output: html_document
---

Read tornado data in and call needed R libraries. 

```{r}
setwd("C:/Users/sgille3/Projects/Course Work/INFO 511 - Analytics/Project 1/tornado/")
library(data.table)
library(ggplot2)
library(dplyr)
library(openintro)
library(choroplethr)
library(choroplethrMaps)

tornado<-fread("tornadoSGEdits.csv")
```

Utilize 'dplyr' package to clean data.

```{r}
dataForPlot <- tornado %>%
                mutate(totalLoss = (money_loss + crop_loss), 
                       totalMortality = (injuries + fatalities),
                       region = tolower(abbr2state(state))) %>%
                select(serialid, state, region, totalMortality, totalLoss, latitude, 
                       longitude, elat, elong, length_miles, width_yards)

head(dataForPlot)
```

Investigate the association between geographical location and mortality.

```{r}
dataMortality <- dataForPlot %>%
                      select(region, totalMortality) %>%
                      group_by(region) %>%
                      summarise(value = sum(totalMortality)) %>%
                      filter(region != '<NA>') %>%
                      arrange(-value)
str(dataForPlot)
head(dataMortality,100)

dataMortality<-as.data.frame(dataMortality)
mortalityPlot<-state_choropleth(dataMortality, title="1950 - 2012 Tornado Morality by State")
```

Investigate the relationship between geographical location and property loss.

```{r}
dataLoss <- dataForPlot %>%
                      select(region, totalLoss) %>%
                      group_by(region) %>%
                      summarise(value = sum(totalLoss)) %>%
                      filter(region != '<NA>') %>%
                      arrange(-value)

dataLoss<-as.data.frame(dataLoss)
lossPlot<-state_choropleth(dataLoss, title="1950 - 2012 Tornado Property Loss by State")
```

Combine plots together to investigate differences in rates of mortality and property loss

```{r}
multiplot(mortalityPlot,lossPlot,cols=2)
```

Combine data for interactive Plotly figure

```{r}
dataForPlotly <- tornado %>%
                    mutate(totalLoss = (money_loss + crop_loss), 
                    totalMortality = (injuries + fatalities),
                    region = abbr2state(state)) %>%
                  select(serialid, state, region, totalMortality, totalLoss, latitude, 
                    longitude, elat, elong, length_miles, width_yards) %>%
                  group_by(state) %>%
                  summarise(mortalitySum = sum(totalMortality), lossSum = sum(totalLoss)) %>%
                  filter(state != '<NA>')
```

Cleaning and plotting final figure for assessment.

```{r}
dataForPlotly$hover <- with(dataForPlotly, paste(state, "<br>", "Economic Loss:", lossSum))

# give state boundaries a white border
l <- list(color = toRGB("black"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('light blue')
)

plot_ly(dataForPlotly, z = mortalitySum, text = hover, locations = state, type = 'choropleth',
        locationmode = 'USA-states', color = mortalitySum, colors = 'Reds',
        marker = list(line = l), colorbar = list(title = "Deaths (thousands)")) %>%
  layout(title = '1950 - 2012 Tornado Morality by State<br>(Hover for Economic Loss in Millions USD)' , geo = g)
```


